\subsection{Broader Impact}
Computer systems affect most parts of our lives. 
Countless examples of the grave consequences 
caused by unreliable or vulnerable software systems
include  severe financial 
damages or even the loss of lives. With the looming advent of 
the internet-of-things, the amount of code  grows at a staggering pace 
demanding for more automated and more sophisticated tools to ensure
software quality.
This project contributes an automated light-weight verification technique that
will help developers write high assurance code.
Further, the project will create a suite of benchmarks to serve as 
baseline for other verification tool developers.  

This project will also engage in efforts to organize a verification competition 
for Java tools. Tool competitions like the \textsf{svcomp}~\cite{svcomp} 
or \textsf{smtcomp}~\cite{smtcomp} have strongly influenced the quality of 
tools in their respective fields, and this project aims at a similar development for Java-based tools.

Although SRI International is a nonprofit research institute rather than a degree-granting educational institution, it makes considerable impact on research and education through its freely distributed tools and  academic visitor programs. The PVS \cite{owre1992pvs}, SAL \cite{bensalem2000overview}, HybridSAL \cite{tiwari2012hybridsal}, and Yices \cite{dutertre2014yices} tools developed at SRI are made freely available. These tools are widely used in academia and industry. In particular, the Yices SMT solver is used as a generic back-end solver. This project will directly contribute to enhancing the reasoning capabilities of Yices, and will have immediate impact on the community of Yices users. Through Visiting Fellow and International Fellow programs, SRI welcomes numerous academic visitors to our Computer Science Laboratory for periods ranging from a few months to several years.  Visitors include professors on sabbatical leave, postdoctoral researchers, and Ph.D. students.  Most of our previous NSF grants have been used to support summer graduate  and postdoctoral students.