
\setcounter{section}{1}
\section{Summary \mytitle}

% Project Summary: The Project Summary consists of an
% overview, a statement on the intellectual merit of the proposed
% activity, and a statement on the broader impacts of the proposed activity.
%
% Please provide between 2 and 6 sets of keywords at the end of the
% overview in the Project Summary. CISE personnel will use this information
% in implementing the merit review process. The keywords should describe the main scientific/engineering areas explored in the proposal. Keywords should be
% prefaced with "Keywords" followed by a colon and each keyword set should be
% separated by semi-colons. Keywords should be of the type used to describe research
% in a journal submission. They should be included at the end of the overview
% in the project summary and might appear, for example, as Keywords: energy-aware
% computing; formal logic; computer graphics; sensor networks; information visualization;
% privacy. "Breakthrough proposals" should have the word "breakthrough" as the
% first keyword in the submitted list of keywords.

% self contained description of activity that would result if proposal
% were funded objective methods employed

{\bf{Overview:}}
%
Writing code is not a manufacturing process, it's not 
always craftsmanship, it's design. Design is where developers add 
value faster than they add cost. However, design is both a complex 
and creative process, which can only be performed by developers after 
much training (over many years). As such, while it would seem a natural 
fit, there is limited work (from a machine learning perspective) 
of how to decouple the design abilities, which allow star developers to 
succeed, from the time it takes to fine tune such abilities. With this 
project, we will address this shortcoming, by applying a data-driven 
approach for program design, called \pdm. 

With \pdm, we will examine ways in which we can get software to borrow 
the same skills and creativity of the formally trained pros, and how 
we can programmatically extend on them to help both traditionally and 
non-traditionally educated programmers to creatively design quality programs. 
\pdm is grounded on the premise that every well-written program, on the 
Web (e.g., GitHub software repositories), represents a concrete example 
of human creativity and design wisdom. As such, it will leverage such a 
readily available wisdom to establish new design interactions by 
identifying, operationalizing, and then manipulating structural 
regularities of (object-oriented) programs, such idioms, coding rules, 
and design patterns.


% Improvements in software verification and automated reasoning techniques are
% leading to a variety of new and highly efficient tools. Recent advances in the
% ability of decision procedures to automatically infer abstractions using
% techniques like Craig interpolation or property-directed reachability have
% paved the way for powerful tools to handle looping and recursive control-flow.
% Although these tools can efficiently complex control-flow, they are usually not
% efficient when reasoning about data structures. Properties about data structures
% generally have to be expressed by universally quantifying over an unbounded number
% of instances which existing techniques do not handle efficiently, if at all.

% To address this limitation, this project proposes a new encoding of
% Java-like languages based on the concept of heap invariants.
% Heap invariants are a light-weight mechanism to abstract a program's
% interactions with data structures that can be combined efficiently
% with existing decision procedures. The project will investigate variations
% of the proposed encoding and build up a set of benchmark problems.

\medskip\noindent{\bf{Keywords}}: program design; design tuning; 
design transfer; software repository mining.   

\medskip\noindent
{\bf{Intellectual Merit:}} 
% 
This project proposes a novel approach for program design and 
evolution. Encouraged by many recent advances in code transplantation, 
data-driven Web design, and software evolution simulation, we 
present \pdm. The key innovations behind this approach include:

\begin{enumerate}
	\item Data-driven program design for (object-oriented) Java programs.
	\item Implementation of semi-automatic program design tuning and evolution.
	\item Learning and operationalization of structural regularities in programs.
\end{enumerate}

\noindent
The project will integrate the developed \pdm technique into an existing 
integrated development environment and then evaluate its effectiveness 
on a series of benchmarks. The goal is to advance the effectiveness and 
performance of Java program design tools and provide a new set of benchmarks 
for a Java program design competition (similar to the software design 
competitions held by TopCoder and Microsoft).


\medskip\noindent
{\bf{Broader Impact:}}
Computer systems affect most parts of our lives. Countless examples of 
the grave consequences caused by poorly designed software systems include 
financial damages or the loss of lives. With the rise of Citizen Programming, 
the amount of code written by opportunistic programmers is growing at an 
astounding rate, demanding for more automated tools to ensure software quality. 
This project contributes a data-driven program design technique that will help 
opportunistic programmers to write code closer to what an experienced programmer 
would write. 

SRI International make considerable contributions to education, research, 
and technology transfer to industry through its freely distributed tools 
and academic visitor programs such as include summer internships for graduate 
students that encourage applications from minority students. 
