\chapter{Summary}{}
\label{sec:intro}

\section*{Overview} % (fold)
\label{sec:overview}

\startWith{w}{riting code} is not a manufacturing process, it's not 
always craftsmanship, it's design. Design is where developers add 
value faster than they add cost. However, design is both a complex 
and creative process, which can only be performed by developers after 
much practice (over many years). In today’s modern society, while the 
number of people with programming ability is growing at an astounding 
rate, the percentage of the population with a well-rounded program 
design experience remains in the single digits. This means that we are 
witnessing a growing gap between need and capacity. One can meet some of 
the need by using configurable programs with canned behavior, but these 
restrict both creativity and problem solving (i.e., capacity) to those 
anticipated by the creators of those programs. Therefore, it is critical 
that we find programmatic ways in which we can help both traditionally 
and non-traditionally educated programmers borrow the same skills and 
creativity of the formally trained pros and still express themselves.

In a new research direction, we term Program Design Mining, we propose  
making some of the abilities, which allow star developers to succeed, 
available to citizen programmers by applying machine learning to program 
design. We will leverage existing design knowledge on the Web (e.g., 
GitHub software repositories) to help program designers. On the Web, 
every well-written program represents a concrete example of human 
creativity and design wisdom. Given these ready available Web data, 
Program Design Mining will identify and then manipulate structural 
regularities \cite{minsky1991law} of (object-oriented) programs, such idioms, 
coding rules, and design patterns. Program Design Mining exploits three machine 
learning applications to enable new interaction mechanisms for program 
design: (1) structured prediction \cite{collins2002discriminative} for 
finding static and semantic correspondences between programs, which can 
be used by programmers to transfer the structural regularities from one 
program into the structure of the other; (2) program induction \cite{lake2015human} 
for learning generative models that could help program designers manipulate 
(and evolve) the transfered structural regularities; and (3) nonparametric 
bayesian analysis \cite{allamanis2014mining} for dealing with incompatibilities 
between structural regularities.

All of these techniques leverage the inherent structure of program designs. 
Well-written (object-oriented) programs, for example, exhibit many structural 
regularities\footnote{These regularities explicitly mold programs with design 
and coding principles that intend to improve their quality} and their 
identification and manipulation improves with experience. Machine learning 
can eliminate the need for gaining design experience, as it allows us to learn 
from existing design knowledge. On the Web data, every Java class is associated 
with an abstract syntax tree. This information can be used along with the change 
history of this class to reason about and manipulate such structural regularities, 
closer to way experienced developers will do. 

With Program Design Mining, we will decouple the ability of designing quality 
programs from the time it takes to fine tune such an ability. Design wisdom 
will not come with many years of practice \cite{weiser1983programming, 
winslow1996programming}. It will come with new design workflows (modeled as code 
transformations) learned from the programs' change history. Developers will be able 
to evolve their programs with sound design and coding principles. At every 
opportunity, they will try a different design alternative, recommended by 
our tools, and experience the results. Program design mining will exploit ready 
available Web data to foster design inspiration when innovation is required, 
and thus reducing the gap between need and capacity.

In what follows, we will cover the intellectual merit of data-driven 
approaches for program design, as well as the broader impact of Program 
Design Mining.

\section*{Intellectual Merit} % (fold)
\label{sec:merit}

This project proposes a novel approach for program design and 
evolution. Encouraged by many recent advances in code transplantation, 
data-driven Web design, and software evolution simulation, we 
present Program Design Mining. The key innovations behind this 
approach include:

\begin{enumerate}
	\item Data-driven program design. Establish new design workflows 
	from existing design knowledge. These workflows allow citizen 
	programmers to automatically transfer the structural regularities 
	from a well-written program into the style and structure of another.  
	\item Semi-automatic program design tuning. Generate structural 
	regularities recommendations by analyzing learned regularities online; 
	i.e., in parallel with structured prediction, which allows the 
	recommendations to adapt to transfer incompatibilities. Programmers 
	can provide feedback on the recommendations, which is taken into 
	account for subsequent recommendations.	
	\item Operationalization and evolution of learned regularities. 
	On the to-be-collected Web data, for example, every Java class is associated 
	with an abstract syntax tree. Along with its change history, we can use 
	this abstract representation to reason about and manipulate programs 
	with \textit{similar} structures.
\end{enumerate}

\section*{Broader Impact} % (fold)
\label{sec:impact}

Computer systems affect most parts of our lives. Countless examples of 
the grave consequences caused by poorly crafted software systems include 
severe financial damages or even the loss of lives. With the looming 
advent of Citizen Programming, the amount of code to be produced by 
opportunistic programmers grows at a staggering pace, demanding for more 
automated tools to ensure software quality. This project contributes a 
data-driven program construction technique that will help people with 
programming ability to write code closer to what an experienced programmer 
would write. Further, the project will create a suite of benchmarks to 
serve as baseline for other automated program construction tools.

SRI International make considerable contributions to education, research, 
and technology transfer to industry through its freely distributed tools 
and academic visitor programs such as include summer internships for graduate 
students that encourage applications from minority students.